---
title: "607 Proj1"
author: "Chris Ayre"
date: "2/22/2019"
output: html_document
---

In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players; 
- Player’s Name
- Player’s State
- Total Number of Points
- Player’s Pre-Rating
- Average Pre Chess Rating of Opponents


load library 

```{r}
require(stringr) 
require(dplyr)
require(tidyr)
require(knitr)
require(readr)
```


load the data

```{r}
tournamentinfo = readLines("/Users/cayre/Google Drive/CUNY SPRING 19/COURSES/data 607/ASSIGNMENTS/tournamentinfo.txt")
```


removing dividers and rows

```{r}
tournamentinfo = str_replace_all(tournamentinfo,"-","")
```

remove empty column

```{r}
tourn = read.delim(textConnection(tournamentinfo),header = FALSE, sep = "|", stringsAsFactors = FALSE)

tourn = tourn[,-11]
```

simplify and add header

```{r}
tournheader = paste(trimws(as.character(tourn[1,])), trimws(as.character(tourn[2,])))

tournheader = make.names(str_sub(tournheader, 1,11))

tourn = as.tbl(tourn[-1:-2,])
names(tourn) = tournheader
```


fix positioning of player id and rank

```{r}
playid = seq(1,nrow(tourn),2)
srank = seq(2,nrow(tourn),2)

tournplayers = tourn[playid,]
stateranks = select(tourn[srank,], State = Pair.Num, Pre.Rating = Player.Name)
```


bind coloumns then display work done

```{r}
tournplayers = bind_cols(tournplayers, stateranks)

tournplayers[1:5,]
```

clean pre rating

```{r}
tournplayers = tournplayers %>% 
  mutate(Pre.Rating = str_replace(Pre.Rating,"\\d+\\s+/ R:\\s+","")) %>% 
  mutate(Pre.Rating = str_replace(Pre.Rating,"^[P].+|>.+","")) %>% 
  mutate(Pre.Rating = as.integer(str_extract(Pre.Rating, "\\d+"))) 
```

clean player name, points and pair

```{r}
tournplayers = tournplayers %>% 
  mutate(Total.Pts = as.numeric(Total.Pts),
         Pair.Num = parse_number(Pair.Num),
         Player.Name = trimws(Player.Name))
```

tidy format

```{r}
tournplayers = tournplayers %>%
  gather("Round","OppPairNum", 4:10) %>% 
  mutate(OppPairNum = parse_number(OppPairNum))
```

join player pre rating

```{r}
pratings = select(tournplayers, Pair.Num,OppPreRating = Pre.Rating) %>% distinct()

tournplayers = left_join(tournplayers, pratings, by = c("OppPairNum" = "Pair.Num"))

```

calculate average opponent rating in final table, preview table

```{r}
tournplayers = tournplayers %>% 
  group_by(Player.Name) %>% 
  mutate(AvgOppPreRating = round(mean(OppPreRating, na.rm=T)))

result = tournplayers %>% 
  select(Player.Name, Player.State = State, Total.Pts, Pre.Rating, AvgOppPreRating) %>% 
  ungroup() %>% 
  distinct()

result[1:10,]
```

export csv

```{r}
write_csv(result, "/Users/cayre/Google Drive/CUNY SPRING 19/COURSES/data 607/ASSIGNMENTS/607-Proj1.csv")
```

